---
title: 'Bios 6301: Assignment 3'
author: "Haoyang Yi"
output:
  pdf_document: default
  html_document: default
---


### Question 1 ###

**15 points**

Write a simulation to calculate the power for the following study
design.  The study has two variables, treatment group and outcome.
There are two treatmenat groups (0, 1) and they should be assigned
randomly with equal probability.  The outcome should be a random normal
variable with a mean of 60 and standard deviation of 20.  If a patient
is in the treatment group, add 5 to the outcome.  5 is the true
treatment effect.  Create a linear model for the outcome by the
treatment group, and extract the p-value (hint: see assigment1).
Test if the p-value is less than or equal to the alpha level, which
should be set to 0.05.

Repeat this procedure 1000 times. The power is calculated by finding
the percentage of times the p-value is less than or equal to the alpha
level.  Use the `set.seed` command so that the professor can reproduce
your results.

1. Find the power when the sample size is 100 patients. (10 points)
```{r}
set.seed(414)
size = 100
times = 1000
res_pvalue = NULL
for(i in 1:times){
group = sample(c(0,1),size,replace = T)
y = rnorm(size,mean = 60,sd = 20)
y[group==1]=y[group==1]+5
mod = lm(y~group)
res_pvalue[i] = coef(summary(mod))[2,4]
}
sum(res_pvalue < 0.05)/times # power when size = 100 


```
1. Find the power when the sample size is 1000 patients. (5 points)
```{r}

set.seed(414)
size = 1000
times = 1000
res_pvalue = NULL
for(i in 1:times){
group = sample(c(0,1),size,replace = T)
y = rnorm(size,mean = 60,sd = 20)
y[group==1]=y[group==1]+5
mod = lm(y~group)
res_pvalue[i] = coef(summary(mod))[2,4]
}
sum(res_pvalue < 0.05)/times # power when size = 1000 


```
### Question 2 ###

**14 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).
Save the `2021/proj_wr21.csv` file in your working directory.  Read
in the data set and remove the first two columns.

1. Show the correlation matrix of this data set. (4 points)
```{r}
library(MASS)
df = read.csv('proj_wr21.csv')
df = df[,-c(1,2)]
cor(df) # correlation matrix
```
1. Generate a data set with 30 rows that has a similar correlation
structure.  Repeat the procedure 1,000 times and return the mean
correlation matrix. (10 points)
```{r}

dfcor = cor(df)
dfcov = var(df)
dfmean = colMeans(df)
rescor = 0
set.seed(414)
# generate data from similar correlation structure
for (i in 1:1000){
    dt = mvrnorm(30,dfmean,dfcov)
    rescor = rescor+cor(dt)/1000 
}
rescor
rescor-cor(df) # The difference between average cor matrix and true cor matrix is small.
```
### Question 3 ###

**21 points**

Here's some code:

```{r}
nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}
as.numeric(sapply(nDist(500), mean)[1])
```

1. What does this do? (3 points)

    ```{r}
    round(sapply(nDist(500), mean), 2)
    ```

  It return 2 digit means of each generated sample from beta,uniform...distributions, like mean of a generated sample with size 500 from beta distribution (5,45)  


1. What about this? (3 points)

    ```{r}
    sort(apply(replicate(20, round(sapply(nDist(10000), mean), 2)), 1, sd))
    ```
    

   It runs the simulation 20 times and calculates the standard deviation of 20 simulated means of each sample from beta,uniform..distributions .   


    In the output above, a small value would indicate that `N=10,000` would provide a sufficent sample size as to estimate the mean of the distribution. Let's say that a value *less than 0.02* is "close enough".

1. For each distribution, estimate the sample size required to simulate the distribution's mean. (15 points)
```{r}
b= nDist(50)
t = data.frame(distribution = names(b),true_mean = c(0.1,16/3,10,1,17/15,1,2,400/150,exp(1/2),32,0,25,0,0.5,1))
n = seq(10,2000,10) # starts with sample size = 10 and set gap = 10
set.seed(414)
diff = matrix(0,length(n),length(names(b)))
for (i in 1:length(names(b))){
    for(j in 1:length(n))
        {diff[j,i]=as.numeric(sapply(nDist(n[j]), mean)[i])-t[i,2] 
}
}
diff = cbind(n,diff)
minval = NULL
for (i in 2:ncol(diff)){
    minval[i-1]=min(which(abs(diff[,i])<=0.02))
}
t$sizerequired = 10+10*(minval-1)
t
```

#Don't worry about being exact. It should already be clear that N < 10,000 for many of the distributions. You don't have to show your work. Put your answer to the right of the vertical bars (`|`) below.

distribution|N
---|---
beta|10
binomial|130
chisquared|130
exponential|90
f|50
gamma|200
geometric|240
hypergeometric|160
lognormal|20
negbinomial|100
normal|80
poisson|360
t|70
uniform|90
weibull|10