---
title: "homework7"
author: "Haoyang Yi"
date: "10/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(lexicon)
```

### Question 1 ###

**21 points**

**45/40, great work! Tianyi**

Use the following code to generate data for patients with repeated measures of A1C (a test for levels of blood glucose).

```{r}
genData <- function(n) {
    if(exists(".Random.seed", envir = .GlobalEnv)) {
        save.seed <- get(".Random.seed", envir= .GlobalEnv)
        on.exit(assign(".Random.seed", save.seed, envir = .GlobalEnv))
    } else {
        on.exit(rm(".Random.seed", envir = .GlobalEnv))
    }
    set.seed(n)
    subj <- ceiling(n / 10)
    id <- sample(subj, n, replace=TRUE)
    times <- as.integer(difftime(as.POSIXct("2005-01-01"), as.POSIXct("2000-01-01"), units='secs'))
    dt <- as.POSIXct(sample(times, n), origin='2000-01-01')
    mu <- runif(subj, 4, 10)
    a1c <- unsplit(mapply(rnorm, tabulate(id), mu, SIMPLIFY=FALSE), id)
    data.frame(id, dt, a1c)
}
x <- genData(500)
```

Perform the following manipulations: (3 points each)

1. Order the data set by `id` and `dt`.
```{r}
x = arrange(x,id,dt)
```
2. For each `id`, determine if there is more than a one year gap in between observations.  Add a new row at the one year mark, with the `a1c` value set to missing.  A two year gap would require two new rows, and so forth.
```{r}
for (i in unique(x$id)){
  for (j in 2:length(x$dt[x$id==i])){
    dtnow = x$dt[x$id==i][j]
    dtbefore = x$dt[x$id==i][j-1]
    ytdiff = as.numeric(dtnow-dtbefore)%/%365
    while (ytdiff>0){
      ytdiff = ytdiff-1
      newrow = data.frame(id=i,dt=dtbefore+years(1),a1c=NA)
      x = rbind(x,newrow)
    }
  }
}
x = arrange(x,id,dt)
```
3. Create a new column `visit`.  For each `id`, add the visit number.  This should be 1 to `n` where `n` is the number of observations for an individual.  This should include the observations created with missing a1c values.

```{r}
x = x %>%
  group_by(id) %>%
  mutate(visit = 1:length(id))
```
4. For each `id`, replace missing values with the mean `a1c` value for that individual.
```{r}
x = x %>%
  group_by(id) %>%
  mutate(a1c = replace_na(a1c,mean(a1c,na.rm = T)))
```
5. Print mean `a1c` for each `id`.
```{r}
x %>%
  group_by(id) %>%
  summarize(mean_a1c=mean(a1c))

```
6. Print total number of visits for each `id`.
```{r}
x %>%
  group_by(id) %>%
  summarize(visits=n())
```
7. Print the observations for `id = 15`.
```{r}
x[x$id==15,]
```
### Question 2 ###

**16 points**

Install the `lexicon` package. Load the `sw_fry_1000` vector, which contains 1,000 common words.

```{r}
data('sw_fry_1000', package = 'lexicon')
head(sw_fry_1000)
```

1. Remove all non-alphabetical characters and make all characters lowercase. Save the result as `a`.

```{r}
a = tolower(gsub("[^A-Za-z ]", "", sw_fry_1000))
```

Use vector `a` for the following questions. (2 points each)

2. How many words contain the string "ar"?
 64
```{r}
length(grep("ar",a))
```

3. Find a six-letter word that starts with "l" and ends with "r".

```{r}
for (i in unique(a)){
if (nchar(i)==6 &startsWith(i,"l")&endsWith(i,"r"))res = i}
res
```

4. Return all words that start with "col" or end with "eck".

```{r}
res = NULL
for (i in unique(a)){
if (startsWith(i,"col")|endsWith(i,"eck"))res = append(res,i)
}
res
```

5. Find the number of words that contain 4 or more adjacent consonants. Assume "y" is always a consonant.
  There are 8 words
```{r}

str_subset(a, "[^aeiou]{4}")


```

6. Return all words with a "q" that isn't followed by a "ui".
   
```{r}
str_subset(a, "q[u][^i]") # q must be followed by a u

```

7. Find all words that contain a "k" followed by another letter. Run the `table` command on the first character following the first "k" of each word.

```{r}
a1 = str_subset(a,"k[^k]")
a2=str_split(a1,"k")
res = NULL
for (i in 1:length(a2)){
res = c(res,a2[[i]][2])  
}
table(substr(res,1,1))
```

8. Remove all vowels. How many character strings are found exactly once?
There are 581 strings found exactly once.
```{r}
a3= gsub("[aeiou]", "", a)
t = table(a3)
count = 0
for (i in 1:length(t)){
if (as.numeric(t[i])==1) count = count+1}
count
```

### Question 3 ###

**3 points**

The first argument to most functions that fit linear models are formulas.  The following example defines the response variable `death` and allows the model to incorporate all other variables as terms. `.` is used to mean all columns not otherwise in the formula.

```{r}

haart_df = read.csv('haart.csv')[,c('death','weight','hemoglobin','cd4baseline')]
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))
```

Now imagine running the above several times, but with a different response and data set each time.  Here's a function:

```{r}
myfun <- function(dat, response) {
  form <- as.formula(response ~ .)
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
```

Unfortunately, it doesn't work. `tryCatch` is "catching" the error so that this file can be knit to PDF.

```{r}
tryCatch(myfun(haart_df, death), error = function(e) e)
```

What do you think is going on?  Consider using `debug` to trace the problem.

The input of the function response cannot be directly assigned as death because it's not a valid value.

**5 bonus points**

Create a working function.
```{r}
myfun <- function(dat, response) {
  y = dat[,response]
  other = dat[,!(colnames(dat) %in% response)]
  coef(summary(glm(y~other[,1]+other[,2]+other[,3], family=binomial(logit))))
}
myfun(haart_df, "death")
````
